---
title: "Table One Demo"
author: "HealthyR"
date: "2023-03-23"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r}
library(tidyverse)
library(finalfit)
```

```{r}
melanoma = boot::melanoma

missing = tibble(time = c(NA,NA),
                 status = c(1,3),
                 sex = c(NA,NA),
                 age = c(65,55),
                 year = c(NA,NA),
                 thickness = c(1.5,NA),
                 ulcer = c(NA, 1))

melanoma_missing = melanoma %>% 
  bind_rows(missing)
```


#Task 1 
Have a quick look at your data 

```{r}
melanoma_missing %>% 
  glimpse()

melanoma_missing %>% 
  ff_glimpse()
```

# Recode
```{r}
# Recode to factors
melanoma_missing = melanoma_missing %>% 
  mutate(sex = factor(sex) %>% fct_recode("Female" = "0",
                                          "Male" = "1"),
         ulcer = factor(ulcer) %>% fct_recode("Absent" = "0",
                                              "Present" = "1"))

melanoma_missing %>% 
  glimpse()

melanoma_missing %>% 
  ff_glimpse()
```

# Label variables
```{r}
melanoma_missing = melanoma_missing %>% 
  mutate(age = ff_label(age, "Age (years)"),
         sex = ff_label(sex, "Sex (at birth)"))
```


# Task 2
Research Question: Is there an association between presence of ulceration and death from melanoma 

Create an example of table 1

```{r}
dependent = "ulcer"
explanatory = c("age", "sex")
melanoma_missing %>% 
  summary_factorlist(dependent, explanatory)
```

# Task 3
Use the missing data dataset - how can you change your table to show/hide missingness

```{r}
dependent = "ulcer"
explanatory = c("age", "sex")
melanoma_missing %>% 
  summary_factorlist(dependent, explanatory, 
                     na_include = TRUE,
                     add_col_totals = TRUE, add_row_totals = TRUE,
                     include_row_missing_col = FALSE)
```

# Task 4
Can you change the labels? Can you add a label for the dependent variable?

```{r}
dependent = "ulcer"
explanatory = c("age", "sex")
melanoma_missing %>% 
  summary_factorlist(dependent, explanatory, 
                     na_include = TRUE, 
                     add_col_totals = TRUE, add_row_totals = TRUE,
                     add_dependent_label = TRUE)
```

# Task 5
Do you include p-values?

```{r}
dependent = "ulcer"
explanatory = c("age", "sex")
melanoma_missing %>% 
  summary_factorlist(dependent, explanatory, 
                     na_include = TRUE, 
                     add_col_totals = TRUE, add_row_totals = TRUE,
                     add_dependent_label = TRUE,
                     p = TRUE)
```

# Task 6

Export

```{r}
dependent = "ulcer"
explanatory = c("age", "sex")
melanoma_missing %>% 
  summary_factorlist(dependent, explanatory, 
                     na_include = TRUE, 
                     add_col_totals = TRUE, add_row_totals = TRUE,
                     add_dependent_label = TRUE,
                     p = TRUE) %>% 
  knitr::kable(row.names = FALSE, align = c("l", "r", "r", "l", "r", "r", "r", "r"))
```
