---
title: "DAG demo"
format: html
editor: visual
---

```{r}
# install.packages("ggdag")
library(ggdag)
library(tidyverse)
```
1.  Basic DAG

node_size text_size

```{r}

dagify(
  outcome ~ exposure) %>% 
    ggdag()

```
2.  We are investigating the causal relationship between social media use and poor mental health outcomes

```{r}
dagify(
  mental_health ~ social_media_use + other_thngs, 
  other_things ~ yet_more_things) %>% 
    ggdag()
```

2.  Add confounders: age and personality type Add mediator: sleep duration

```{r}
dagify(
  mental_health ~ social_media_use + age + personality + sleep_duration,
  sleep_duration ~ social_media_use,
  social_media_use ~ age + personality
  ) %>% 
    ggdag()
```

3.  Coordinates

manually set coords time ordered
```{r}
coords <- list(
  x = c(mental_health = 5, social_media_use = 0, age = 2.5, personality = 2.5, sleep_duration = 2.5),
  y = c(mental_health = 1, social_media_use = 1, age = 4, personality = 6, sleep_duration = -1)
)
dagify(
  mental_health ~ social_media_use + age + personality + sleep_duration,
  sleep_duration ~ social_media_use,
  social_media_use ~ age + personality,
  
  coords = time_ordered_coords()
  ) %>% 
    ggdag()
```

```{r}
coords <- list(
  x = c(mental_health = 5, social_media_use = 0, age = 2.5, personality = 2.5, sleep_duration = 2.5),
  y = c(mental_health = 1, social_media_use = 1, age = 4, personality = 6, sleep_duration = -1)
)
dagify(
  mental_health ~ social_media_use + age + personality + sleep_duration,
  sleep_duration ~ social_media_use,
  social_media_use ~ age + personality,
  
  coords = coords,
  
  exposure = "social_media_use",
  outcome = "mental_health",
  
  labels = c(mental_health = "mental health",
             social_media_use = "social media use",
             age = "age",
             personality = "personality",
             sleep_duration = "sleep duration")
  ) %>% 
    ggdag_status(use_labels = "label",
                 text = FALSE)
  
```

4.  Add labels and status

Within dagify set exposure = "", outcome = "", labels = c( ="")

try ggdag() and ggdag_status()

5.  Add theme

```{r}
dagify(
  mental_health ~ social_media_use + age + personality + sleep_duration,
  sleep_duration ~ social_media_use,
  social_media_use ~ age + personality,
  
  coords = coords,
  
  exposure = "social_media_use",
  outcome = "mental_health",
  
  labels = c(mental_health = "mental health",
             social_media_use = "social media use",
             age = "age",
             personality = "personality",
             sleep_duration = "sleep duration")
  ) %>% 
    ggdag_status(use_labels = "label",
                 text = FALSE)+
  theme_dag()
```
```{r}
dagify(
  mental_health ~ social_media_use + age + personality + sleep_duration,
  sleep_duration ~ social_media_use,
  social_media_use ~ age + personality,
  
  coords = coords,
  
  exposure = "social_media_use",
  outcome = "mental_health",
  
  labels = c(mental_health = "mental health",
             social_media_use = "social media use",
             age = "age",
             personality = "personality",
             sleep_duration = "sleep duration")
  ) %>% 
  ggdag_adjustment_set()
```


6.  Dag Usability

open paths (adjust_for) minimal adjustment set

```{r}
dagify(
  poor_mental_health ~ social_media + age  + personality + sleep,
  social_media ~ age + personality,
  sleep ~ social_media, 
  exposure = "social_media",
  outcome = "poor_mental_health",
  coords = coords,
  labels = c(
    poor_mental_health = "Poor Mental Health",
    social_media = "Social Media Usage",
    age = "Age (years)",
    personality = "Personality Type",
    sleep = "Sleep (hours per night)")) %>% 
  as_tidy_dagitty() %>% 
  ggplot(aes(x, y, xend = xend, yend = yend)) +
  geom_dag_edges() + 
  geom_dag_point()+
  #geom_dag_text()+
  geom_text(aes(label=label, x=x, y=y), color = "blue")
  
```

7.  Additional customisations in ggplot

e.g geom_text themes colour palettes

```{r}

dagify(
  poor_mental_health ~ social_media + age  + personality + sleep,
  social_media ~ age + personality,
  sleep ~ social_media, 
  exposure = "social_media",
  outcome = "poor_mental_health",
  coords = coords,
  labels = c(
    poor_mental_health = "Poor Mental Health",
    social_media = "Social Media Usage",
    age = "Age (years)",
    personality = "Personality Type",
    sleep = "Sleep (hours per night)")) %>% 
  as_tidy_dagitty() %>% 

  ggplot(aes(x, y, xend = xend, yend = yend)) +
  geom_dag_edges() + 
  geom_dag_point()+
  geom_dag_text()+

  

```

### ANSWERS

basic -

```{r}
set.seed(1)
dagify(
  poor_mental_health ~ social_media + age  + personality + sleep,
  social_media ~ age + personality,
  sleep ~ social_media, 
  exposure = "social_media",
  outcome = "poor_mental_health") %>% 
  ggdag(node_size = 20,
        text_size = 3)
```

add coordinates -

```{r}
coords <- list(
  x = c(social_media = 1, poor_mental_health = 10, age = 3, personality = 5, sleep = 7),
  y = c(social_media = 5, poor_mental_health = 5, age = 1, personality = 1, sleep = 7)
)
```

```{r}
basic_dag = dagify(
  poor_mental_health ~ social_media + age  + personality + sleep,
  social_media ~ age + personality,
  sleep ~ social_media, 
  exposure = "social_media",
  outcome = "poor_mental_health",
  coords = coords)

basic_dag %>% 
  ggdag()


basic_dag = dagify(
  poor_mental_health ~ social_media + age  + personality + sleep,
  social_media ~ age + personality,
  sleep ~ social_media, 
  exposure = "social_media",
  outcome = "poor_mental_health",
  coords = time_ordered_coords())

basic_dag %>% 
  ggdag()
```

add labels and status -

```{r}

basic_dag = dagify(
  poor_mental_health ~ social_media + age  + personality + sleep,
  social_media ~ age + personality,
  sleep ~ social_media, 
  exposure = "social_media",
  outcome = "poor_mental_health",
  coords = coords,
  labels = c(
    poor_mental_health = "Poor Mental Health",
    social_media = "Social Media Usage",
    age = "Age (years)",
    personality = "Personality Type",
    sleep = "Sleep (hours per night)"))

basic_dag %>% 
  ggdag(use_labels = "label", text = F)
  
basic_dag %>% 
  ggdag_status(use_labels = "label", text = F)
```

add theme -

```{r}
basic_dag %>% 
  ggdag(use_labels = "label", text = F) +
  theme_dag()


basic_dag %>% 
  ggdag(use_labels = "label", text = F) +
  theme_dag_grid()

basic_dag %>% 
  ggdag(use_labels = "label", text = F) +
  theme_classic()
```

DAG fetures

```{r}
# what are the open pathways 
basic_dag %>% 
  ggdag_paths()

basic_dag %>% 
  ggdag_paths(adjust_for = "age")



# what adjustment set do I need to use?
basic_dag %>% 
  ggdag_adjustment_set(shadow = TRUE)

basic_dag %>% 
  dag_adjustment_sets()


basic_dag %>% 
  ggdag_parents("poor_mental_health")

basic_dag %>% 
  ggdag_dseparated()
```

Additional customization

```{r}
dagify(
  poor_mental_health ~ social_media + age  + personality + sleep,
  social_media ~ age + personality,
  sleep ~ social_media, 
  exposure = "social_media",
  outcome = "poor_mental_health",
  coords = coords,
  labels = c(
    poor_mental_health = "Poor Mental Health",
    social_media = "Social Media Usage",
    age = "Age (years)",
    personality = "Personality Type",
    sleep = "Sleep (hours per night)")) %>% 
  as_tidy_dagitty() %>% 
  
  
  mutate(extra = case_when(name == "age" ~ "Confounder",
                           name == "personality" ~ "Confounder",
                           name == "poor_mental_health" ~ "Outcome",
                           name == "sleep" ~ "Mediator",
                           name == "social_media" ~ "Exposure")) %>% 

  ggplot(aes(x, y, xend = xend, yend = yend)) +
  geom_dag_edges() + 
  geom_dag_point(size = 20, aes(colour = extra))+
  # geom_dag_text(size = 2.4)+
  theme_void()+
  scale_color_brewer(palette = "Set3")+
  # theme(legend.position = "none")+
  geom_text(aes(label = label, x =x, y = y ), size = 3)
```

edit colours in ggplot: https://stackoverflow.com/questions/71441472/change-color-for-specific-nodes-in-ggdag
