---
title: "Markdown and exporting your plots/tables"
author: 'Riinu Pius'
date: today
format:
  html: 
    code-fold: true
    code-tools: true
  pdf: default
  docx: default
---
​
# Data and set-up
​
Loading the packages:
​
```{r message=FALSE, warning=FALSE}
​
library(tidyverse)
library(knitr)
library(gapminder)
​
```
​
We are using the gapminder dataset. This is a document I'm working in, this text will be nicely printed.
​
```{r}
​
mydata = gapminder %>% 
  mutate(gdpPercap = round(gdpPercap)) %>% 
  mutate(pop_millions = round(pop/1e6, 1)) %>% 
  select(-pop)
​
```
​
A random sample of rows in the dataset:
​
```{r}
​
mydata %>% 
  sample_n(10) %>% 
  kable()
​
```
​
Number of variables: `r mydata %>% ncol()`.
​
Total number of observations: `r mydata %>% nrow()`.
​
# Plotting
​
Results can be seen in \@fig-rplot-1
​
```{r}
#| label: fig-rplot
#| fig-cap: "gapminder GDP per capita vs life expectancy globally"
#| fig-subcap: 
#| - "Original plot"
#| - "New plot"
#| layout-ncol: 2
​
​
mydata %>% 
  filter(year == 2007) %>% 
  ggplot(aes(x      = gdpPercap/1000, #divide by 1000 to tidy the x-axis
             y      = lifeExp,
             colour = continent,
             size   = pop_millions)) +
  geom_point(shape = 1) +
  theme_bw()
​
mydata %>% 
  filter(year == 2007) %>% 
  ggplot(aes(x      = gdpPercap/1000, #divide by 1000 to tidy the x-axis
             y      = lifeExp,
             colour = continent,
             size   = pop_millions)) +
  geom_point(shape = 1) +
  theme_bw() +
  scale_colour_brewer()
```
​
```{=html}
<!--
:::{.content-visible when-profile="teacher"}
-->
```
::: {.content-visible when-format="pdf"}
# Solutions
​
This how we solve this. Don't show it to students!!!! You're doing a great job.
:::
