---
title: "Tidymodels"
format: html
editor: visual
---

# Tidymodels Demo

Tidymodels is a collection of packages for modelling in R. One of the main developers is Max Kuhn, who developed the Caret package, and tidymodels uses a lot of this functionality, in combination with other packages for modelling. Tidymodels utilises tidyverse-style programming.

### Helpful links:

https://www.tmwr.org/ - tidymodels book

https://www.tidymodels.org/find/recipes/ - preprocessing options

https://www.tmwr.org/pre-proc-table.html - example preprocessing recipe

https://www.tidymodels.org/find/parsnip/ - model specification options

## Data

```{r}
library(tidyverse)
library(tidymodels)

diabetes = read_csv("diabetes_binary_health_indicators.csv") %>% 
  sample_n(5000) 

diabetes = diabetes %>% 
  mutate(across(c("Diabetes_binary", "HighBP", "HighChol", "CholCheck", "Smoker", "Stroke",
                  "HeartDiseaseorAttack", "PhysActivity", "Fruits", "Veggies", 
                  "HvyAlcoholConsump", "AnyHealthcare", "NoDocbcCost", "DiffWalk", "Sex"), as.factor)) %>% 
  mutate(added = 1,
         subjid = 1:n(),
         Diabetes_binary = case_when(Diabetes_binary == "1" ~ "Yes",
                                     Diabetes_binary == "0" ~ "No",
                                     .default = NA) %>% fct_relevel("Yes"))


diabetes %>% 
  glimpse()

diabetes = diabetes %>% 
  janitor::clean_names()

```

## Data Split

```{r}
split = diabetes %>% initial_split(prop = 0.8)
diabetes_train = training(split)
diabetes_test = testing(split)
```

## Recipes

https://www.tidymodels.org/find/recipes/

-   remove columns with zero varience
-   create dummy variable
-   change role - update_role()

```{r}
diabetes_recipe = 
  recipe(diabetes_binary ~ ., data = diabetes_train) %>% 
  step_zv() %>% 
  step_
```

## Specifications

https://www.tidymodels.org/find/parsnip/

Start with decision tree

Change to logistic regression?

```{r}

diabetes_tree_spec = decision_tree(cost_complexity = 0.0001) %>%
  set_engine("rpart") %>% 
  set_mode("classification") 

```

## Workflows

Use workflow()

```{r}
diabetes_workflow = 
  workflow() %>% 
  add_recipe(_) %>% 
  add_model(_)

diabetes_fit = 
  fit(_)
```

## Data Metrics

roc_auc() roc_curve() conf_mat()

```{r}
# predict with augment()
augment(diabetes_fit, new_data = diabetes_train)



```

## Evaluate Test Set

last_fit() Using workflow and data split

```{r}

```

# ANSWERS

## Recipes

https://www.tidymodels.org/find/recipes/

-   remove columns with zero varience

-   create dummy variable

-   normalise

-   change role

```{r}
diabetes_recipe = 
  recipe(diabetes_binary ~ ., data = diabetes_train) %>% 
  step_zv() %>% 
  step_dummy() %>% 
  update_role(subjid, new_role = "id") 

diabetes_recipe %>% prep() %>% juice()
```

## Specifications

https://www.tidymodels.org/find/parsnip/

Start with decision tree

Change to logistic regression?

```{r}
diabetes_tree_spec = decision_tree(cost_complexity = 0.0001) %>%
  set_engine("rpart") %>% 
  set_mode("classification") 

diabetes_lr_spec = logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")
```

## Workflows

Use workflow()

```{r}
diabetes_workflow = 
  workflow() %>% 
  add_recipe(diabetes_recipe) %>% 
  add_model(diabetes_tree_spec)

diabetes_fit = 
  fit(diabetes_workflow, diabetes_train)
```

## Data Metrics

```{r}
# predict with augment()
augment(diabetes_fit, new_data = diabetes_train)%>%
  roc_auc(truth = diabetes_binary, .pred_Yes)

augment(diabetes_fit, new_data = diabetes_train)%>%
  roc_curve(truth = diabetes_binary, .pred_Yes) %>% 
  autoplot()

augment(diabetes_fit, new_data = diabetes_train)%>%
  conf_mat(truth = diabetes_binary, .pred_class) %>% 
  autoplot(type = "heatmap")
```

## Evaluate

```{r}
diabetes_fit_final = last_fit(diabetes_workflow, split) 

collect_metrics(diabetes_fit_final)
```

## Resamples

```{r}
diabetes_folds = bootstraps(diabetes_train)


resample_diabetes = 
  fit_resamples(diabetes_workflow, diabetes_folds, metrics = metric_set(accuracy, roc_auc, brier_class), 
                         control = control_resamples(save_pred = TRUE))

resample_diabetes %>% 
  collect_predictions() %>% 
  roc_auc(truth = diabetes_binary, .pred_Yes) 

resample_diabetes %>% 
  collect_predictions() %>% 
  roc_curve(truth = diabetes_binary, .pred_Yes) %>%
  autoplot()
```
